Verifying Spatial Properties of Array Computations

### 2017-04-17 oopsla17/evaluation.tex
bounded specs
added a table of multiple use of mulOp; included smagorinsky example commented because it seems to be no longer the case under our new model...
limitations & jacobi

>>>>

\textbf{Bounded} specifications occured with \num{\overallatMost}
\texttt{atMost} bounds and \num{\overallatLeast} \texttt{atLeast}, the latter of which were
always also paired with an upper bound.

% \begin{figure}[t]\begin{minted}[fontsize=\scriptsize,breakindent=0em,linenos=false,xleftmargin=0em,breakafter=)]{fortran}
% !=stencil readOnce,(forward(depth=1,dim=3,nonpointed))*(backward(depth=1,dim=1))*(backward(depth=1,dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1))*(pointed(dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1,nonpointed))*(backward(depth=1,dim=2,nonpointed))+(forward(depth=1,dim=3))*(backward(depth=1,dim=2))*(pointed(dim=1))+(backward(depth=1,dim=1))*(backward(depth=1,dim=2,nonpointed))*(pointed(dim=3))+(backward(depth=1,dim=1,nonpointed))*(backward(depth=1,dim=2))*(pointed(dim=3))::x
% \end{minted}
% \caption{Complex specification inferred from
%   \textbf{UM}\label{fig:smagorinsky}}
% \vspace{-1em}
% \end{figure}

% The single- and multi-action classes represent more complex stencils
% with a real possibility for programmer error. As an extreme example, the Unified
% Model has an implementation of the Smagorinsky subgrid-scale model for
% calculating turbulence on which our inference yields 39 specs from 340
% lines of code. This is a large reduction given
% the complexity of the algorithm.  We show one example
% in~\Cref{fig:smagorinsky} which specifies the access pattern to
% a 3-dimensional array (which we have renamed to
% \texttt{x}) arising from a kernel of 93 lines of code involving 142
% array subscripts. The specification was the most complex
% seen in our corpus, yet it still represents a significant abstraction
% of the spatial behaviour given size and complexity of the kernel it describes.

We measured the frequency at which individual specifications involved
multiple occurences of \term{*}
 % and \term{+}:
\begin{center}
\vspace{0.25em}
\setlength{\tabcolsep}{0.57em}
{\small{
\hspace{-1em}\begin{tabular}{c|cccccccc}
& 0 & 1 & 2 & 3 & 4 & 5 & 6  \\ \hline
\term{*} & \num{\overallmulOpszero} & \num{\overallmulOpsone} & \num{\overallmulOpstwo} & \num{\overallmulOpsthree} & \num{\overallmulOpsfour} & \num{0} & \num{\overallmulOpssix}
\end{tabular}}}
\vspace{0.2em}
\end{center}
\noindent
% \eg{}, roughly half the specifications did not involve $\term{*}$,
% a quarter use one $\term{*}$ and just under a quarter
% use two $\term{*}$ operators.

\paragraph{Limitations}

There were various reasons why we did not infer
specifications on every looped array computation: \\
\noindent
1) \textbf{Non-subset induction variables} occur when the
induction variables on the RHS are not a subset of those in the LHS. These
cases are not stencils by our definition. The degenerate case of this
is to have only constant indices on the LHS. We see lots of examples
of this in loops as accumulators \eg{} computing the sum over an array; \\
\noindent
2) \textbf{Derived induction variables} where an
index (\mintinline{fortran}{x}) is derived from an
induction variable (\mintinline{fortran}{i}) as in
\mintinline{fortran}{x = len - i};  \\
\noindent
3) \textbf{Inconsistent induction dimensions} occur when
an induction variable is used to specify more than one array dimension
on the RHS or multiple induction variables are used for the same
dimension on the RHS. These are common in matrix operations such as
LU-decomposition with assignments such as
\mintinline{fortran}{a(l) = a(l) - a(m) * b(l, m)}.

\subsection{Detecting errors in the 2-D Jacobi iteration}
\label{sec:jacobi}
One common example of a stencil computation is the two-dimensional
Jacobi iteration that repeatedly goes through each cell in a matrix
and computes the average value of the four adjacent cells. The kernel
is given by:
\begin{minted}{fortran}
  a(i,j) = (a(i-1,j)+a(i+1,j)+a(i,j+1)+a(i,j-1))/4
\end{minted}
We infer a precise specification of its shape as:
\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  != stencil pointed(dim=1)*centered(depth=1,dim=2,nonpointed)+pointed(dim=2)*centered(depth=1,dim=1,nonpointed)::a
\end{minted}
%
To test our implementation,
we examined whether programmer errors would be detected by replacing
the array index offsets with $-1$, $0$, or $1$ and running our
verification algorithm. Our checking procedure correctly reported a verification
failure in each of 6,537 permutations corresponding to an error.
The iteration computes the average of four adjacent
cells so $24$ ($4$ factorial) of the possible array index perturbations
are correct, all of which are accepted by our checker.
